Efficient and Effective Metric Space Indexing for Complete Record Linkage

### 2017-11-20 paper.tex
Update on Overleaf.

Record linkage, also known as entity resolution, data matching and
duplicate detection~\cite{Chr12}, is the process of identifying and
matching records that refer to the same real-world entities within or
across databases. The entities to be linked are often people (such as
patients in hospital or customers in business databases), but record
linkage can also be applied to link consumer products or bibliographic
records~\cite{Chr12}. Record linkage is commonly challenged by the lack
of unique entity identifiers (keys) in the databases to be linked, which
prevents the use of a database join. Instead, the linkage of records
requires the comparison of the common attributes (or fields) that are available
in the datasets to be linked. For datasets that contain
information about individuals, these attributes include the names,
address, dates of birth, and so on, of individuals.

@@

\begin{figure}[!t]
  \centering
  %\scalebox{1.0}[1.0]
  \includegraphics[width=1.0\textwidth]{figures/linkage-process}
  \caption{Overview of the steps of the traditional record linkage
           process (left-hand side) and our proposed metric space
           indexing based approach (right-hand side), as described in
           Sect.~\ref{sec-intro}, where records from two databases,
           $\mathbf{D}_A$ and $\mathbf{D}_B$, are being linked.}
           \label{fig-rl-process}
\end{figure}

@@

% do we need a section on preliminaries, notation, etc?
% ideally techniques are described formally, possibly via pseudo-code
% algorithmic descriptions

@@

% peter: We need one paragraph describing the overall 'approach' /
% methodology, as well as the aim, then a short description of
% each method/technique used

% would be good to have a figure if possible illustrating these
% approaches

% Graham - should it be complete/incomplete rather than exact/inexact below?
We explore the efficacy of several similarity search algorithms in
record linkage: one popular inexact algorithm, LSH-minhash, and one exact
method, M-tree~\cite{paolociaccia2m}. We also use a simple brute force
(nested-loop) exact technique as a baseline for comparison, although
this approach can only feasibly be applied to the smallest of our
datasets. All experiments have a number of parameters with which we can
configure the search space and algorithm behaviour, as described in
Sect.~\ref{sec-exp}. A \emph{distance threshold}, $d$, specifies a
maximum distance, equivalent to a minimum similarity, for two records to
be classified as a link (i.e.\ referring to the same entity).

@@

% Needed? Apart from anything else, haven't introduced the datasets yet.
% Each of our datasets contains ground truth with which we can establish the records that are true links and those that are not. Clearly this approach can only be applied to relatively small datasets and we restrict its usage to the Cora dataset as a baseline for the other experiments.

\smallskip
\textbf{LSH-MinHash:}
The LSH-minhash algorithm used has three tuning parameters:
\emph{shingle size}, $l_{ss}$, the \emph{band size}, $l_{bs}$, and the
\emph{number of bands}, $l_{nb}$.
These are explained in the following algorithm description which follows a standard LSH approach [CITATION].
Using this approach, each record drawn from a set of records is placed in a \emph{LSH-MinHash} data structure.
To do this, for each record, firstly all the fields of the record are turned into strings and 
concatenated. Next the strings are \emph{shingled} into a set of
n-grams (sub-strings of length $n$) according to the $l_{ss}$
parameter (i.e.\ $n = l_{ss}$). Next a set of randomly, 
deterministically generated hash functions are applied to each of the
sets of n-grams and the smallest (the minhash) of each application is
added to a signature. The number of hashes used and thus the size of
the signature is set to $l_{nb} \times l_{bs}$.
Lastly the signature is divided into $l_{nb}$ segments and the elements from the band are hashed again to create a number of keys.
The original record is added to a hashmap associated with each of the keys. 
We we will evaluate the performance of LSH based blocking with regard to the three parameters $l_{ss}$, $l_{bs}$ and $l_{nb}$ in
Sect.~\ref{sec-exp}.
To perform linkage the following procedure is followed: for each query value is hashed using the procedure described above and the union of set of of values found for all hashes in the hashmap are returned. These are then be filtered by performing distance calculations on the results to eliminate false positives.

@@

% subsection

%\subsection

%\subsection{Analysis and Limitations}
%\label{sec-analysis}

%maybe a complexity analysis? maybe an analysis of expected 
%linkage quality?

@@

We now describe the data sets we use in our evaluation, the setup we
employed to evaluate our proposed metric indexing approach and
compare it with traditional blocking techniques, and we then present
and discuss our results.

@@

@@

\begin{table}[t]
\caption{Parameter settings used for the different data sets used in
   the experiments,with $d$ the distance threshold used, $t_b$ the
   M-tree branching factor, $l_{ss}$, $l_{sb}$ and $l_{nb}$ the LSH
   shingle size, and the size and number of bands, respectively.
 \label{table-parameters}
  \centering
  \begin{scriptsize}
  %\addtolength{\tabcolsep}{-0.5pt}
  \begin{tabular}{cccccccc}
  \hline\noalign{\smallskip}
  Data set~ & $d$ & $t_b$ & $l_{ss}$ & $l_{bs}$ & $l_{nb}$ & 
    $f_{ki}$ & $f_{ks}$ \\
  name(s) & & & & & & & \\
  \noalign{\smallskip} \hline \noalign{\smallskip}
  CORA & ~$[0,5,10,\ldots,95,100]$~ & ~$[5,10,25,50]$~ & ~$2$~ &
    ~$[2,5,10]$~ & ~$[2,5,10]$~ & ~$[30,35,40]$~ & ~$[30,35,40]$ \\
  NCVR  &  \\
  Isle of Skye &  \\
  Kilmarnock  &  \\
  \noalign{\smallskip} \hline
  \end{tabular}
  \end{scriptsize}
\end{table}

>>>>

Record linkage, also known as entity resolution, data matching and
duplicate detection~\cite{Chr12}, is the process of identifying and
matching records that refer to the same real-world entities within or
across datasets. The entities to be linked are often people (such as
patients in hospital or customers in business datasets), but record
linkage can also be applied to link consumer products or bibliographic
records~\cite{Chr12}. Record linkage is commonly challenged by the lack
of unique entity identifiers (keys) in the datasets to be linked, which
prevents the use of a database join. Instead, the linkage of records
requires the comparison of the common attributes (or fields) that are available
in the datasets to be linked. For datasets that contain
information about individuals, these attributes include the names,
address, dates of birth, and so on, of individuals.

@@

\begin{figure}[!t]
  \centering
  %\scalebox{1.0}[1.0]
  \includegraphics[width=1.0\textwidth]{figures/linkage-process}
  \caption{Overview of the steps of the traditional record linkage
           process (left-hand side) and our proposed metric space
           indexing based approach (right-hand side), as described in
           Sect.~\ref{sec-intro}, where records from two datasets,
           $\mathbf{D}_A$ and $\mathbf{D}_B$, are being linked.}
           \label{fig-rl-process}
\end{figure}

@@

@@

We explore the efficacy of several similarity search algorithms for
record linkage: traditional blocking, one popular incomplete algorithm, LSH-minhash, and one complete
method, M-tree~\cite{paolociaccia2m}. We also use a simple brute force
(nested-loop) exact technique as a baseline for comparison, although
this approach can only feasibly be applied to the smallest of our
datasets. All experiments have a number of parameters with which we can
configure the search space and algorithm behaviour, as described in
Sect.~\ref{sec-exp}. A \emph{distance threshold}, $d$, specifies a
maximum distance, equivalent to a minimum similarity, for two records to
be classified as a link (i.e.\ referring to the same entity).

@@

\smallskip
\textbf{Traditional  blocking:}
In the traditional blocking technique records are blocked using hand selected keys as described in \cite{Chr12b}. The keys are chosen with the intention of giving the best possible results for the dataset. Multiple blocking keys are commonly used to address the problem of missing and erroneous values preventing matches from being found.

\smallskip
\textbf{LSH-MinHash:}
The LSH-minhash algorithm used has three tuning parameters:
\emph{shingle size}, $l_{ss}$, the \emph{band size}, $l_{bs}$, and the
\emph{number of bands}, $l_{nb}$.
These are explained in the following algorithm description which follows a standard LSH approach [CITATION].
Using this approach, each record drawn from a set of records is placed in a \emph{LSH-MinHash} data structure.
To do this, for each record, firstly all the fields of the record are turned into strings and 
concatenated. Next the strings are \emph{shingled} into a set of
n-grams (sub-strings of length $n$) according to the $l_{ss}$
parameter (i.e.\ $n = l_{ss}$). Next a set of randomly, 
deterministically generated hash functions are applied to each of the
sets of n-grams and the smallest (the minhash) of each application is
added to a signature. The number of hashes used and thus the size of
the signature is set to $l_{nb} \times l_{bs}$.
Lastly the signature is divided into $l_{nb}$ segments and the elements from the band are hashed again to create a number of keys.
The original record is added to a hashmap associated with each of the keys. 
We we will evaluate the performance of LSH based blocking with regard to the three parameters $l_{ss}$, $l_{bs}$ and $l_{nb}$ in
Sect.~\ref{sec-exp}.

To perform linkage the following procedure is followed: for each query, the value is hashed using the procedure described above and the union of the set of values found for all hashes in the hashmap are returned. These are then be filtered by performing distance calculations on the results to eliminate false positives.

@@

@@

We now describe the data sets we use in our evaluation, the setup
employed to evaluate our proposed metric indexing approach and
compare it with traditional blocking techniques, and we then present
and discuss our results.

@@

%The experiments for each technique have been run within a generic Java framework which permits the actual linkage method used to be supplied. This ensures that the algorithms are consistently executed and that true comparisons of the their behaviour can be reported. It supplies each algorithm with the source records to be matched using an \emph{addSourceRecord} method which takes a record as a parameter. For each of the records to be matched a  \emph{getCandidateMatches} method is called which is expected to return a list of the linked records from the sources. Each of our datasets contains ground truth with which we can establish the records that are true links and those that are not (discussed below). From this ground truth the framework can report the number of links make, linkage quality etc.

@@

\begin{table}[t]
\caption{Parameter settings used for the different data sets used in
   the experiments,with $d$ the distance threshold used, $t_b$ the
   M-tree branching factor, $l_{ss}$, $l_{sb}$ and $l_{nb}$ the LSH
   shingle size, and the size and number of bands, respectively.}
 \label{table-parameters}
  \centering
  \begin{scriptsize}
  %\addtolength{\tabcolsep}{-0.5pt}
  \begin{tabular}{cccccccc}
  \hline\noalign{\smallskip}
  Data set~ & $d$ & $t_b$ & $l_{ss}$ & $l_{bs}$ & $l_{nb}$ & 
    $f_{ki}$ & $f_{ks}$ \\
  name(s) & & & & & & & \\
  \noalign{\smallskip} \hline \noalign{\smallskip}
  CORA & ~$[0,5,10,\ldots,95,100]$~ & ~$[5,10,25,50]$~ & ~$2$~ &
    ~$[2,5,10]$~ & ~$[2,5,10]$~ & ~$[30,35,40]$~ & ~$[30,35,40]$ \\
  NCVR  &  \\
  Isle of Skye &  \\
  Kilmarnock  &  \\
  \noalign{\smallskip} \hline
  \end{tabular}
  \end{scriptsize}
\end{table}
